The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-10-05 paper/ms.tex
update model 3

For planets detected in binary systems,
\begin{equation}
N_{{\rm det},d}
= 
N_d \Gamma_{t,d} (f_{d1,g}f_{d1,c} + f_{\rm desired}w_{d2}f_{d2,g}f_{d2,c}).
\label{eq:N_det_d_model2}
\end{equation}
The factor $f_{\rm desired}$, as previously discussed, accounts for the 
fraction of secondaries not of the desired {\it stellar} type. The above 
counts the total number of detected planets around stars of the desired type, 
though all these planets have diluted radii.
The detection and geometric transit probabilities for a given system are
\begin{align}
f_{d1,c} &= (1+\gamma_R)^{-3} \\
f_{d1,g} &= f_{s,g} \\
f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-5/\alpha} 
\label{eq:f_d2c}\\
f_{d2,g} &= f_{s,g}q^{2/3},
\end{align}
where typically $\alpha=3.5$.
To derive Eq.~\ref{eq:f_d2c}, one must note that for fixed period, planet 
radius, and signal to noise, 
$$d_{\rm det}\propto \mathcal{D} R_2^{-2} L_d^{1/2} T_{\rm dur}^{1/2}.$$
For a typical binary drawn from the local stellar neighborhood, $\gamma_R 
\approx 0.2$ (see Fig.~\ref{fig:gammaR_distribn_mag_limited}), and there is an 
interesting asymmetry.
One finds the completeness for planets which orbit the primary star, 
$f_{d1,c}$ to be
\begin{equation}
f_{d1,c} \sim 70\% \quad(\alpha=3.5, \gamma_R=0.2).
\end{equation}
However, for planets which orbit the secondary star, the completeness is
\begin{equation}
f_{d2,c} \sim 7\% \quad(\alpha=3.5, \gamma_R=0.2).
\end{equation}
This is the obvious statement that controlling for all other variables, 
in the typical binary system planets around secondaries are more difficult to 
detect than those around primaries.

@@

Instead, we run a Monte Carlo simulation in which we fix the planet and 
primary star properties, draw the binary mass ratio from 
Eq.~\ref{eq:pdf_q_mag_limited}, 
and take the luminosity as $L\propto M^{3.5}$. We find a mean occurrence 
correction of 
\begin{equation}
\langle X_\Gamma \rangle =1.268\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
\label{eq:X_Gamma_model2_lower}
\end{equation}
and a median that is the same to within one percent.
This value is comparable to what one would guess using the analytic results 
for $\langle N_d \rangle / N_s$ to substitute $\beta$, and $\gamma_R$ to 
compute $f_{d1,c}$.

@@

Keep the same stars as in Sec.~\ref{sec:model_2}.
Now let the planets have a power-law radius distribution, ${\rm 
d}(N_p/N_\star)/{\rm 
d}R_p\propto R_p^\delta$. 
We want to know the effects of binarity on the apparent distribution of 
planetary radii.
Specifically, the average number of planets in a given size range per star of 
the desired type in this universe is
\begin{equation}
\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} = 
\Gamma_0 \int_{R_p^{\rm \ell}}^{R_p^{\rm u}} R_p^\delta \,{\rm d}R_p,
\end{equation}
for $\Gamma_0$ one of $\{\Gamma_{t,s}, \Gamma_{t,d}, w_{d2}\Gamma_{t,d}\}$
corresponding to single star systems, the primaries of double star systems, 
and the secondaries of double star systems.

@@

\appendix
\begin{figure}[!t]
    \begin{center}
        \includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
    \end{center}
    \caption{Empirical fit to main sequence dwarf mass luminosity data 
    compiled 
        from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' 
        is a 
        least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' 
        is 
        to 
        data above that, and below the Kraft break ($1.2M_\odot$).
        The $L(M)$ relation taken for numerics is the maximum 
        of the two fits. For analytics, we assume $L\propto M^{3.5}$.
    }
    \label{fig:mass_luminosity}
\end{figure}

>>>>

For planets detected in binary systems,
\begin{equation}
N_{{\rm det},d}
= 
N_d \Gamma_{t,d} (f_{d1,g}f_{d1,c} + f_{\rm desired}w_{d2}f_{d2,g}f_{d2,c}).
\label{eq:N_det_d_model2}
\end{equation}
The factor $f_{\rm desired}$, as previously discussed, accounts for the 
fraction of secondaries not of the desired {\it stellar} type. The above 
counts the total number of detected planets around stars of the desired type, 
though all these planets have diluted radii.
The detection and geometric transit probabilities for a given system are
\begin{align}
f_{d1,c} &= (1+\gamma_R)^{-3} 
\label{eq:f_d1c}
\\
f_{d1,g} &= f_{s,g} \\
f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-5/\alpha} 
\label{eq:f_d2c}\\
f_{d2,g} &= f_{s,g}q^{2/3},
\label{eq:f_d2g}
\end{align}
where typically $\alpha=3.5$.
To derive Eq.~\ref{eq:f_d2c}, one must note that for fixed period, planet 
radius, and signal to noise, 
$$d_{\rm det}\propto \mathcal{D} R_2^{-2} L_d^{1/2} T_{\rm dur}^{1/2}.$$
For a typical binary drawn from the local stellar neighborhood, $\gamma_R 
\approx 0.2$ (see Fig.~\ref{fig:gammaR_distribn_mag_limited}), and there is an 
interesting asymmetry.
One finds the completeness for planets which orbit the primary star, 
$f_{d1,c}$ to be
\begin{equation}
f_{d1,c} \sim 70\% \quad(\alpha=3.5, \gamma_R=0.2).
\end{equation}
However, for planets which orbit the secondary star, the completeness is
\begin{equation}
f_{d2,c} \sim 7\% \quad(\alpha=3.5, \gamma_R=0.2).
\end{equation}
This is the obvious statement that controlling for all other variables, 
in the typical binary system planets around secondaries are more difficult to 
detect than those around primaries.

@@

Instead, we run a Monte Carlo simulation in which we fix the planet and 
primary star properties, draw the binary mass ratio from 
Eq.~\ref{eq:pdf_q_mag_limited}, 
and take the luminosity as $L\propto M^{3.5}$. We find a mean occurrence 
correction of 
\begin{equation}
\langle X_\Gamma \rangle =1.273\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
\label{eq:X_Gamma_model2_lower}
\end{equation}
and a median that is the same to within one percent.
This value is comparable to what one would guess using the analytic results 
for $\langle N_d \rangle / N_s$ to substitute $\beta$, and $\gamma_R$ to 
compute $f_{d1,c}$.

@@

For this model, we keep the same stars as in Sec.~\ref{sec:model_2}, but the 
planets have a power-law radius distribution, ${\rm 
d}(N_p/N_\star)/{\rm 
d}R_p\propto R_p^\delta$. 
We want to know the effects of binarity on the apparent distribution of 
planetary radii.
Specifically, the average number of planets in a given size range per star of 
the desired type in this universe is
\begin{align}
\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} 
&= 
\int_{R_p^{\rm \ell}}^{R_p^{\rm u}} \Gamma(R_p) \,{\rm d}R_p
\\
&=
\Gamma_0 \int_{R_p^{\rm \ell}}^{R_p^{\rm u}} R_p^\delta \,{\rm d}R_p,
\end{align}
for $\Gamma_0$ one of $\{\Gamma_{t,s}, \Gamma_{t,d}, w_{d2}\Gamma_{t,d}\}$
corresponding to single star systems, the primaries of double star systems, 
and the secondaries of double star systems.
More strictly, $\Gamma$ should now be referred to as an occurrence rate 
density (it always has been, but in previous models the planet-radius 
dependence was a delta function).

The observer now calculates occurrence rates slightly differently.
The original survey was of stars with $m<m_{\rm min}$, and the
planet detections are limited by the achievable signal-to-noise from this 
sample.
For a given planet radius, the observer must now count $N_\star (R_p)$ stars 
that they think were searchable.
In reality, some of these stars will be binaries:
\begin{equation}
N_\star(R_p) = N_s(R_p) + N_d(R_p),
\end{equation}
where the number of selected stars in single and double star systems is 
analogous to those of Sec.~\ref{sec:model_1}, except they scale as the 
sixth power of planet radius.
The completeness should now be thought of as
\begin{equation}
{\rm completeness} = \frac{{\rm actual\ number\ of\ searchable\ stars}}{{\rm 
assumed\ number\ of\ searchable\ stars}}.
\end{equation}
If the completeness is calculated in this way, 
Equations~\ref{eq:f_d1c} through~\ref{eq:f_d2g} remain the same.

@@

\newpage
\appendix
\begin{figure}[!t]
    \begin{center}
        \includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
    \end{center}
    \caption{Empirical fit to main sequence dwarf mass luminosity data 
    compiled 
        from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' 
        is a 
        least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' 
        is 
        to 
        data above that, and below the Kraft break ($1.2M_\odot$).
        The $L(M)$ relation taken for numerics is the maximum 
        of the two fits. For analytics, we assume $L\propto M^{3.5}$.
    }
    \label{fig:mass_luminosity}
\end{figure}
